IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-03-02 paper/proofs.tex
Move full proof to appendix.

\begin{proof}[Proof of Theorem~\ref{thm:iso-tsni}]
  We proof the theorem by induction on the length of the derivation sequence in~\eqref{eq:tsni-lhs}.
  The base case for derivations
  of length 0 is trivial, allowing
  us to simple chose $|ic2'=ic2|$.  In the step case,
  we split the derivation sequence from~\eqref{eq:tsni-lhs} of length $n+1$ as follows:
  \[
  |ic1| |.->| |ic1''| |.->|^n |ic1'|
  \]
  for some configuration |ic1''|.  By Lemma~\ref{lemma:rr-tsni-general} to
  the first step and
  the induction hypothesis to the rest of the derivation sequence,
  we directly get the desired property.
  % long version
%  We proof the theorem by induction on the length of the derivation sequence in~\eqref{eq:tsni-lhs}.
%  The base case for derivations
%  of length 0 is trivial, allowing
%  us to simple chose $|ic2'=ic2|$.  In the step case, we assume
%  the theorem holds for derivation sequences of length up to $n$, and show that it also
%  holds for those of length $n+1$.  We split the derivation sequence from~\eqref{eq:tsni-lhs} as follows:
%  \[
%  |ic1| |.->| |ic1''| |.->|^n |ic1'|
%  \]
%  for some configuration |ic1''|.  By Lemma~\ref{lemma:rr-tsni-general}, we get
%  |ic''| with
%  \begin{equation} \label{eq:tsni-proof-1}
%  |ic1''| \approx_{|il|} |ic2''|
%  \qquad \text{and} \qquad
%  |ic2| |.->|^* |ic2''|
%  \end{equation}
%  Applying the induction hypothesis to
%  $|ic1''| |.->|^n |ic1'|$, we get |ic2'| with
%  \begin{equation} \label{eq:tsni-proof-2}
%  |ic1'| \approx_{|il|} |ic2'|
%  \qquad \text{and} \qquad
%  |ic2''| |.->|^* |ic2'|
%  \end{equation}
%  Stitching together the derivation sequences from~\eqref{eq:tsni-proof-1} and~\eqref{eq:tsni-proof-2} directly gives
%  us the right-hand side of the implication in the TSNI
%  definition~\eqref{eq:tsni-rhs}, which concludes the proof.
\end{proof}
\begin{proof}[Proof of Lemma~\ref{lemma:rr-tsni-general}]
  First, we observe there must be at least one task in |ic1|, otherwise
  it could not take a step.  Thus, |ic1| is of the form
  |iconf iS1 (it1, its1)|.  Consider two cases:
  \begin{itemize}
    \item $|erase il it1|=|bullet|$.
    By the definition of |erasef il|, we know that |il canFlowTo lcurr|
    where |lcurr| is the label of |it1|.
    In this case, we do not need to take a step for
    |ic2|, because |ic2'=ic2| will already be |il|-equivalent to |ic1'|.
    To see that, note that the tasks |its1| in |ic1| are left in the
    same order and unmodified (the scheduling policy only
    modifies the first task). The task |it1| either
    gets dropped (by \textsc{I-noStep}), or
    transforms into a task |it1'| as well as potentially spawning a new
    task |it1''|.  Since both |it1'| and |it1''| have a label that is
    at least as high as the label of |it1| (can be seen
    by inspecting all reduction rules), they will get filtered
    by |erasef il| in |ic1'|.  Therefore, the |il|-equivalence of the
    task list is guaranteed.
    Lets consider the possible changes to |iS1|:
    Only three reduction interact with |iS1|,
    thus it suffices to consider these cases:
    \begin{description}
      \item[Case \textsc{I-send}]
      A new message triple with label |il'| gets added to the message
      queue of the receiving thread.  However, since |lcurr canFlowTo il'|,
      the triple will get erased.
      \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
      In this case, only the queue of
      task |it1| can change, which gets erased.
    \end{description}
    This ensures that $|ic1'|\approx_{|il|}|ic2'|=|ic2|$.
    %    \alphacondition{We need all scheduling policies to not change the order
    %      of any tasks (except for the first one).  Newly spawned task can appear
    %      anywhere in the list.}
    \item $|erase il it1|\neq|bullet|$.
    Here, there must be a corresponding
    task |it2| in |ic2|,
    such that |it1=it2| (otherwise |ic1| and
    |ic2| could not be |il|-equivalent).
    However, |it2| might not be at the beginning of the task list yet, but
    all tasks occurring before it must get erased by |erasef il|.
    In |ic2|, we can first take some number of steps until |it2| moves
    to the front of the list.
    This is the case regardless of any additional side conditions $P$ on
    rules, because for all of these tasks, it can either take an actual
    step, or it gets dropped by \textsc{I-noStep} (which is always
    possible, as there cannot be an additional side condition on this
    rule).  All tasks that didn't get dropped are still at a label
    that isn't below |il| and thus get erased.

    %    \alphacondition{The scheduling policy must eventually let any task in
    %      the task list evaluate.  In particular, it cannot get stuck when the
    %      first task gets stuck, or keep executing a small number of tasks
    %      exclusively forever (e.g. just execute the first task all the time
    %      if it gets into an infinite loop).}

    Therefore, after |ic2| has potentially executed some number of steps
    to arrive at |ic2''|, we are now in the situation where $|ic1|\approx_{|il|}|ic2''|$, and the first tasks |it1| and |it2|,
    respectively, don't get erased and are thus equivalent.
    The task |it2| can now take exactly the same step as |it1|;  this
    is true even with arbitrary additional premises $P$ that follow
    the condition in Definition~\ref{def:restricted}, since those
    predicates only depend on |erase il ic1|, which is equivalent
    to |erase il ic2|, and thus those predicates evaluate in the same way.
    Thus, we only
    need to argue that the potential differences in |iS1| and |iS2| cannot
    have an influence on the execution (and we know |iS1| and |iS2| are
    |il|-equivalent).
    Again, only sending and receiving will depend on, or change |iS|,
    so we consider all these cases.
    \begin{description}
      \item[Case \textsc{I-send}]
      Here, the task |it2| will send the same message to the same
      receiver queue. This
      queue is either completely erased, or it is |il|-equivalent.  In both
      cases, |il|-equivalence of |iS1'| and |iS2'| is preserved.
      \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
      When the tasks are receiving a message, then by the reduction rules
      we know that they first filter the queue by the label
      |lcurr| of |it1|.  We
      also know that the queues are equivalent when filtered by the less
      restrictive label |il|, thus the messages received (or dropped) from the
      queue are equivalent.
    \end{description}
  \end{itemize}
\end{proof}
The proof of Theorem~\ref{thm:iso-tini} about TINI proceeds largely
in the same fashion as for Theorem~\ref{thm:iso-tsni} (TSNI), except that
some cases are simpler due to the fact termination is not an issue
(cf.~\cite{lio} for a TINI proof of a similar system).

>>>>

\begin{proof}[Proof of Theorem~\ref{thm:iso-tsni}]
  We proof the theorem by induction on the length of the derivation sequence in~\eqref{eq:tsni-lhs}.
  The base case for derivations
  of length 0 is trivial, allowing
  us to simple chose $|ic2'=ic2|$.  In the step case,
  we split the derivation sequence from~\eqref{eq:tsni-lhs} of length $n+1$ as follows:
  \[
  |ic1| |.->| |ic1''| |.->|^n |ic1'|
  \]
  for some configuration |ic1''|.  By Lemma~\ref{lemma:rr-tsni-general} to
  the first step and
  the induction hypothesis to the rest of the derivation sequence,
  we directly get the desired property.
  % long version
%  We proof the theorem by induction on the length of the derivation sequence in~\eqref{eq:tsni-lhs}.
%  The base case for derivations
%  of length 0 is trivial, allowing
%  us to simple chose $|ic2'=ic2|$.  In the step case, we assume
%  the theorem holds for derivation sequences of length up to $n$, and show that it also
%  holds for those of length $n+1$.  We split the derivation sequence from~\eqref{eq:tsni-lhs} as follows:
%  \[
%  |ic1| |.->| |ic1''| |.->|^n |ic1'|
%  \]
%  for some configuration |ic1''|.  By Lemma~\ref{lemma:rr-tsni-general}, we get
%  |ic''| with
%  \begin{equation} \label{eq:tsni-proof-1}
%  |ic1''| \approx_{|il|} |ic2''|
%  \qquad \text{and} \qquad
%  |ic2| |.->|^* |ic2''|
%  \end{equation}
%  Applying the induction hypothesis to
%  $|ic1''| |.->|^n |ic1'|$, we get |ic2'| with
%  \begin{equation} \label{eq:tsni-proof-2}
%  |ic1'| \approx_{|il|} |ic2'|
%  \qquad \text{and} \qquad
%  |ic2''| |.->|^* |ic2'|
%  \end{equation}
%  Stitching together the derivation sequences from~\eqref{eq:tsni-proof-1} and~\eqref{eq:tsni-proof-2} directly gives
%  us the right-hand side of the implication in the TSNI
%  definition~\eqref{eq:tsni-rhs}, which concludes the proof.
\end{proof}
\begin{proof}[Proof sketch of Lemma~\ref{lemma:rr-tsni-general}]
  For space reasons, we only sketch the structure of the proof here,
  and refer the interested reader to the appendix where we give the
  full proof.
  The proof distinguishes two cases:
  \begin{itemize}
    \item If the task |it1| executed in |ic1| is erased, then there is no
    need to take a step in |ic2|, as the computation in |ic1| does
    not have any visible effects after erasure.
    \item Otherwise, there must be a corresponding task |it2| in |ic2|, and
    it can take exactly the same reduction step as the task in |ic1|.
    The only difficulty is that there might be secret tasks executing
    before |it2|, but it can be shown that these do not influence the
    computation after erasure.
  \end{itemize}
\end{proof}
The proof of Theorem~\ref{thm:iso-tini} about TINI proceeds largely
in the same fashion as for Theorem~\ref{thm:iso-tsni} (TSNI), except that
some cases are simpler due to the fact termination is not an issue
(cf.~\cite{lio} for a TINI proof of a similar system).
